---
title: "Custom geometry: district-level NTL with custom boundaries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Custom geometry: district-level NTL with custom boundaries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

`lightson` accepts any `sf` polygon layer. The workflow below starts with the
bundled LGD state boundaries, then swaps in district-level GeoJSONs from
[bharatviz](https://github.com/saketlab/bharatviz).

## State-level workflow (bundled boundaries)

```{r state-boundaries}
library(lightson)
library(sf)

states <- get_india_admin("state")
nrow(states)
head(states[, c("state_name", "state_code")])
```

Download Bhuvan WMS rasters for a subset of years. `bhuvan_raster()` derives
the bounding box from the geometry:

```{r state-rasters}
rasters <- bhuvan_raster(states, years = 2020:2022)
names(rasters)
rasters[["2022"]]
```

Bhuvan WMS returns RGB visualisation tiles, not physical radiance.
`bhuvan_raster()` converts to single-band luminance via
`lum = 0.2126*R + 0.7152*G + 0.0722*B` (ITU-R BT.709). Values are
comparable across years for trend analysis but are not in nW/cm^2/sr.

```{r state-panel}
panel <- extract_panel(rasters, states, id_col = "state_name")
head(panel)
```

```{r state-trend, fig.width = 7, fig.height = 4}
library(ggplot2)
sel <- c("Bihar", "Maharashtra", "Tamil Nadu", "Uttar Pradesh")
plot_ntl_trend(
  panel[panel$region_id %in% sel, ],
  region   = sel,
  caption  = "Source: ISRO Bhuvan NTL portal"
)
```

```{r state-map, fig.width = 7, fig.height = 6}
plot_ntl_map(
  rasters[["2022"]],
  polygons = states,
  title    = "India nighttime lights, 2022"
)
```


## District-level workflow

```{r}
BHARATVIZ <- "https://bharatviz.org"

districts <- tryCatch(
  sf::read_sf(file.path(BHARATVIZ, "India-bhuvan-districts.geojson")),
  error = function(e) NULL
)
if (is.null(districts)) {
  message("bharatviz not reachable -- skipping district examples")
  knitr::opts_chunk$set(eval = FALSE)
}
nrow(districts)
```

```{r download-rasters}
rasters_dist <- bhuvan_raster(districts, years = 2018:2023)
names(rasters_dist)
```

```{r extract-panel}
panel_dist <- extract_panel(rasters_dist, districts, id_col = "district_name")
head(panel_dist)
```

```{r plot-trend, fig.width = 7, fig.height = 4}
plot_ntl_trend(panel_dist, region = c("Mumbai", "Delhi", "Chennai", "Bengaluru"))
```

```{r plot-map, fig.width = 7, fig.height = 6}
plot_ntl_map(rasters_dist[["2022"]], polygons = districts)
```

### LGD boundaries

```{r lgd}
lgd <- sf::read_sf(file.path(BHARATVIZ, "India_LGD_districts.geojson"))
panel_lgd <- extract_panel(rasters_dist, lgd, id_col = "district_name")
head(panel_lgd)
```

### Historical census boundaries

```{r historical}
dist_2011 <- sf::read_sf(file.path(BHARATVIZ, "India-2011-districts.geojson"))
panel_2011 <- extract_panel(rasters_dist, dist_2011)
head(panel_2011)
```

`extract_panel()` picks the first character column as `region_id` automatically,
or pass `id_col` explicitly.

---

## VIIRS path

To use NASA VIIRS rasters, supply an [Earthdata token](https://urs.earthdata.nasa.gov/):

```{r viirs}
token <- earthdata_token()
rasters_viirs <- ntl_download("viirs", region = districts, years = 2018:2023, token = token)
if (length(rasters_viirs) > 0) {
  panel_viirs <- extract_panel(rasters_viirs, districts, id_col = "district_name")
  head(panel_viirs)
}
```
