---
title: "India state-level NTL with Bhuvan"
description: >
  Pull pre-aggregated state-level radiance from the Bhuvan NTL REST API
  and plot trends across Indian states.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{India state-level NTL with Bhuvan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

ISRO's Bhuvan NTL portal provides pre-aggregated mean radiance for all Indian
states from 2012 to 2024, and WMS raster tiles for any custom region, with no
authentication required.

## All states, all years

```{r all-states}
library(lightson)

df <- bhuvan_stats()
dim(df)
head(df)
```

## Single state time series

```{r single-state}
bihar <- bhuvan_stats(state = "BIHAR")
bihar
```

```{r plot-bihar, fig.width = 7, fig.height = 4}
library(ggplot2)
plot_ntl_trend(bihar,
  region = "Bihar",
  caption = "Source: ISRO Bhuvan NTL portal"
)
```

Radiance nearly 10x from 2012 to 2024.

## Compare states

```{r compare-states, fig.width = 7, fig.height = 4}
states <- c("Bihar", "Maharashtra", "Tamil Nadu", "Uttar Pradesh", "Kerala")
plot_ntl_trend(
  df[df$state %in% states, ],
  region   = states,
  subtitle = "Bhuvan NTL luminance (ITU-R BT.709)",
  caption  = "Source: ISRO Bhuvan NTL portal"
)
```

## Nighttime India from space, 2012 to 2024


```{r india-map-data}
states_sf <- get_india_admin("state")

years_sel <- c(2012, 2015, 2018, 2021, 2024)
rasters <- bhuvan_raster("IND", years = years_sel, bbox_pad = c(0, 0, 0, 3.5))

if (length(rasters) == 0) stop("No rasters downloaded. Check network / Bhuvan WMS.")
```

```{r india-maps, fig.width = 11, fig.height = 8}
plot_ntl_panel(
  rasters,
  polygons = states_sf,
  title    = "India at night, 2012 to 2024",
  subtitle = "Bhuvan NTL luminance (ITU-R BT.709). Boundaries: LGD/SoI.",
  caption  = "Source: ISRO Bhuvan NTL portal",
  ncol     = 3L
)
```

Each panel uses the same luminance scale anchored at zero, so brightness is
comparable across years. The sharpest gains are in Bihar, UP, and
Odisha after 2015, broadly tracking the [Saubhagya electrification scheme](https://powermin.gov.in/en/content/saubhagya)
(2017-2019).

## Filter by year range

```{r filter-years}
post_2015 <- bhuvan_stats(years = 2015:2024)
nrow(post_2015)
```
