---
title: "Choosing and comparing nighttime-light sources"
description: >
  Compare Bhuvan and NASA Black Marble coverage, units, authentication,
  and appropriate India workflows.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choosing and comparing nighttime-light sources}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
bhuvan_up <- tryCatch(lightson::bhuvan_stats(years = 2024), error = function(e) NULL)
has_earthdata <- !is.na(Sys.getenv("EARTHDATA_TOKEN", unset = NA)) ||
  !is.na(Sys.getenv("EARTHDATA_USERNAME", unset = NA))
live_ok <- !is.null(bhuvan_up) && has_earthdata

knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = live_ok)
```

```{r offline-notice, echo = FALSE, results = "asis", eval = !live_ok}
cat("_This vignette needs both the ISRO Bhuvan API and a NASA Earthdata token, and at least one was unavailable when this vignette was built, so its live-data examples were skipped. Run this vignette locally (with `EARTHDATA_TOKEN` set) to see the output._")
```

`lightson` exposes related but non-interchangeable products. Choose a
source based on the period, update frequency, measurement units, and intended
analysis. Do not append their raw values into a single time series.

## At a glance

| Source/product | Coverage | Cadence | Approx. resolution | Values | Authentication |
|---|---:|---:|---:|---|---|
| ISRO Bhuvan NTL | 2012--2024 | Annual | 500 m | RGB-derived BT.709 luminance | None |
| NASA VNP46A4 | 2012--2025 | Annual | 500 m | Radiance, nW/cm^2/sr | Earthdata token |
| NASA VNP46A3 | 2012--present | Monthly | 500 m | Radiance, nW/cm^2/sr | Earthdata token |
| NASA VNP46A2 | 2012--near-present | Daily | 500 m | Radiance, nW/cm^2/sr | Earthdata token |

Availability can lag acquisition. In particular, completed annual composites
are normally published well into the following year. Check the provider when
reproducibility depends on an exact latest date.

## Bhuvan: simple India annual data

Bhuvan is convenient when an unauthenticated India-only workflow matters. Its
state statistics endpoint returns physical radiance statistics, but its WMS
serves rendered RGB imagery. `bhuvan_raster()` therefore converts WMS colours
to dimensionless luminance; it cannot recover the original radiance.

```{r bhuvan}
library(lightson)

state_stats <- bhuvan_stats(years = 2012:2024)
bhuvan_maps <- bhuvan_raster("IND",
  years = 2020:2022,
  bbox_pad = c(0, 0, 0, 3.5)
)
```

Bhuvan uses VNP46A4 Collection 1 for 2012-2023 and Collection 2 from 2024.
The processing change creates a discontinuity, so a 2023-2024 difference
must not be interpreted directly as growth. For a consistent calibrated
2012-2025 annual panel, prefer NASA Collection 2 for every year.

## NASA Black Marble: annual, monthly, and daily

All three NASA products use the same Earthdata bearer token:

```{r black-marble}
token <- earthdata_token()

annual <- ntl_download(
  "IND",
  years = 2020:2022, token = token,
  product = "annual"
)

monthly_2025 <- ntl_download(
  "IND",
  dates = seq(as.Date("2025-01-01"), as.Date("2025-02-01"), by = "month"),
  token = token, product = "monthly"
)

recent_days <- ntl_download(
  "IND",
  token = token, product = "daily",
  dates = seq(as.Date("2025-11-01"), as.Date("2025-11-02"), by = "day")
)
```

Daily downloads default to `quality = "high"`, retaining mandatory QA flag
zero. This rejects cloud contamination and other poor retrievals but can leave
missing pixels. Use `quality = "all"` only when you intend to inspect and
apply the quality bands yourself. Daily data are near-current rather than a
true real-time feed.

Use annual data for long-run development panels, monthly data for seasonal or
short-run economic change, and daily data for narrowly dated events such as
power outages. Aggregating QA-filtered daily observations into a monthly
statistic is not automatically equivalent to the official VNP46A3 composite.

## Comparison

Use the same polygons and aggregation statistic for both sources. Here the
comparison is restricted to 2020--2022 so it does not cross Bhuvan's 2024
collection break.

```{r comparison}
districts <- get_india_admin("district")

viirs_panel <- extract_panel(annual[names(annual) %in% 2020:2022],
  districts,
  id_col = "district_name"
)
viirs_panel$source <- "NASA VNP46A4 C2"
viirs_panel$units <- "nW/cm^2/sr"

bhuvan_panel <- extract_panel(bhuvan_maps[names(bhuvan_maps) %in% 2020:2022],
  districts,
  id_col = "district_name"
)
bhuvan_panel$source <- "ISRO Bhuvan WMS"
bhuvan_panel$units <- "BT.709 luminance"
```

Raw levels cannot be compared because the units differ. A useful diagnostic is
to express each district and source relative to its own 2020 value. An index of
120 then means that the source reports a 20% increase since 2020; it does not
mean that the two sensors observed the same absolute brightness.

```{r indexed-comparison}
comparison <- rbind(viirs_panel, bhuvan_panel)
comparison <- ntl_index(
  comparison,
  baseline = 2020,
  group_cols = c("source", "region_id"),
  index_col = "index_2020"
)

selected <- c("Mumbai", "Patna", "Bengaluru", "Lucknow")
plot_data <- comparison[comparison$region_id %in% selected, ]

library(ggplot2)
ggplot(plot_data, aes(year, index_2020, colour = source)) +
  geom_hline(yintercept = 100, colour = "grey60", linetype = 2) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~region_id) +
  labs(
    x = NULL, y = "Index (2020 = 100)",
    title = "Within-source nighttime-light change",
    subtitle = "Index trajectories are comparable; raw levels are not",
    colour = NULL
  ) +
  theme_ntl() +
  theme(legend.position = "bottom")
```

Check whether conclusions depend on the source instead of relying only on the
figure. The following produces one correlation per district across the three
overlap years and ranks districts with the weakest agreement:

```{r agreement}
agreement <- ntl_source_agreement(viirs_panel, bhuvan_panel)
head(agreement, 10)
```

With only three overlap observations this correlation is a diagnostic, not a
validation statistic. 

## Analyse monthly seasonality

`extract_panel()` recognises ISO-date raster names and adds a `date` column.
This makes the monthly product directly usable as a temporal panel:

```{r monthly-analysis}
monthly_panel <- extract_panel(monthly_2025, districts,
  id_col = "district_name"
)

monthly_plot <- monthly_panel[monthly_panel$region_id %in% selected, ]
ggplot(monthly_plot, aes(date, mean_radiance, colour = region_id)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.8) +
  labs(
    x = NULL, y = "Mean radiance (nW/cm^2/sr)",
    title = "Monthly nighttime lights in 2025",
    colour = NULL
  ) +
  theme_ntl() +
  theme(legend.position = "bottom")
```

For formal seasonality analysis, download multiple complete years and include
month effects. A single year cannot distinguish a recurring seasonal pattern
from a one-off shock.

## Analyse a daily event window

Daily values are most informative alongside coverage. A sudden radiance fall
with a simultaneous collapse in `n_pixels` is more likely missing retrievals
than a power outage.

```{r daily-analysis}
daily_panel <- extract_panel(recent_days, districts,
  id_col = "district_name"
)
event <- daily_panel[daily_panel$region_id == "Delhi", ]

old_par <- par(mfrow = c(2, 1), mar = c(3, 4, 2, 1))
plot(event$date, event$mean_radiance,
  type = "b",
  xlab = "", ylab = "Mean radiance",
  main = "Daily light and valid-pixel coverage"
)
plot(event$date, event$n_pixels,
  type = "h",
  xlab = "Date", ylab = "Valid pixels"
)
par(old_par)
```

For an outage study, use pre-event and post-event windows, compare affected
districts with unaffected controls, and record the exact QA rule. Do not fill
cloudy daily observations with zero.

For most India economic panels, the preferred core series is VNP46A4
Collection 2 from 2012 onward. Bhuvan remains useful for unauthenticated India
exploration and for its pre-aggregated state statistics.
